Automatic Matching of Bullet Lands

### 2016-10-07 imaging-paper.Rnw
Include reference to x3pr

Throughout this paper, we work with images from the James Hamby Consecutively Rifled Ruger Barrel Study~\citep{hamby:2009}. Ten consecutively rifled Ruger P-85 pistol barrels were obtained from the manufacturer and fired to produce 20 known test bullets and 15 unknown bullets for comparison. 
3D topographical images of each bullet were obtained using a NanoFocus lens at 20x magnification and made publicly available on the NIST Ballistics Database Project\footnote{\url{http://www.nist.gov/forensics/ballisticsdb/hamby-consecutively-rifled-barrels.cfm}} in a format called x3p (XML 3-D Surface Profile). The x3p format conforms to the ISO5436-2 standard\footnote{\url{http://sourceforge.net/p/open-gps/mwiki/X3p/}}, implemented to provide a simple and standard conforming way to exchange 2D and 3D profile data. It was adopted by the OpenFMC (Open Forensic Metrology Consortium\footnote{\url{http://www.openfmc.org/}}), a group of academic, industry, and government firearm forensics researchers whose aim is to establish best practices for researchers using metrology in forensic science.

@@

A procedure for bullet matching using the BulletTrax3D system is described in \citet{roberge:2006}. This study is based on another set of ten consecutively rifled barrels; matches are identified based on a bullet-to-bullet correlation score. The authors state that this process `could be automated', but no implementation of the algorithm is available. Modern automated techniques using 3D images have also been conducted by \citet{riva:2014}, but this work applies to striae found on cartridge cases and not bullets.

>>>>

Throughout this paper, we work with images from the James Hamby Consecutively Rifled Ruger Barrel Study~\citep{hamby:2009}. Ten consecutively rifled Ruger P-85 pistol barrels were obtained from the manufacturer and fired to produce 20 known test bullets and 15 unknown bullets for comparison. 
3D topographical images of each bullet were obtained using a NanoFocus lens at 20x magnification and made publicly available on the NIST Ballistics Database Project\footnote{\url{http://www.nist.gov/forensics/ballisticsdb/hamby-consecutively-rifled-barrels.cfm}} in a format called x3p (XML 3-D Surface Profile). The x3p format conforms to the ISO5436-2 standard\footnote{\url{http://sourceforge.net/p/open-gps/mwiki/X3p/}}, implemented to provide a simple and standard conforming way to exchange 2D and 3D profile data. It was adopted by the OpenFMC (Open Forensic Metrology Consortium\footnote{\url{http://www.openfmc.org/}}), a group of academic, industry, and government firearm forensics researchers whose aim is to establish best practices for researchers using metrology in forensic science. An open-source package for working with x3p files has been written in R \citep{R} by Dr. Nicholas Petraco which we build upon in the development of our algorithm \citep{x3pr}. This package enables the direct reading and manipulation of x3p files.

@@

A procedure for bullet matching using the BulletTrax3D system is described in \citet{roberge:2006}. This study is based on another set of ten consecutively rifled barrels; matches are identified based on a bullet-to-bullet correlation score. The authors state that this process `could be automated', but no implementation of the algorithm is available. Modern automated techniques using 3D images have also been conducted by \citet{riva:2014}, but this work applies to striae found on cartridge cases and not bullets. Other work on 3D images has been conducted by \citet{petraco:2012}, which also describes work carried out on cartridge cases, as well as screwdriver striation patterns, with promising results. We aim to use similar ideas of automated 3D image analysis, and apply this work to bullet surfaces with the modern tools enabled by the R statistical software.
